|

### 2020-12-01 paper/v1/broadband.speed.covid.Rmd
Paper edits, esp tests descriptives in Methods section.

<!-- our approach -->
Using the UK as a case study, this paper aims to understand how the quality and
reliability of internet service, as reflected in *experienced* internet speeds,
may reinforce or redress the spatial and social dimensions of the digital division
exposed by the pandemic.
To do so, we employ volunteered geographic data on individual broadband speed tests
and state-of-the-art time-series clustering methods to create clusters of UK local
authorities with similar temporal signatures of experienced internet speeds.
We then associate these clusters of local authorities with their socioeconomic
and geographic characteristics to explore how they overlap with or diverge from
the existing economic and digital geography of the UK. **need to add urban-rural to analysis**
Our analysis enables us to better understand how the spatial and social distribution
of both occupations and online accessibility intersect to enable or hinder the
practice of telecommuting at a time of extreme demand.
We will also consider what lessons can be learned from this time for a future where
telecommuting is likely to remain a common means of accessing work, and broadband
services and infrastructure must be fit for purpose.
**LET'S LEAVE IT FOR NOW, BUT I THINK WE CAN CRYSTALISE MORE THE RQ**

@@

The structure of this paper is as follows.
First we review the literature on telecommuting and digital divides to better
understand the origins of the new digital divide revealed by the pandemic and its
impact on the economic resilience of different places.
We then describe our data and methodology.
Our results section first offers classification of how internet services vary
across the UK local authorities and then assesses whether these clusters replicate
or repudiate other socio-economic and geographic patterns of economic resilience.

@@

<!-- Who works from home-->
In this analysis, the terms ‘telecommuting’ and ‘working from home’ are used
interchangeably, as most remote labour during the Covid-19 crisis was carried out
in the homes of individual employees rather than any other location [@eurofound2020].
However, it should be noted that previous research has explored how telecommuting
can occur in other places, including satellite offices or on public transport
[@felstead2012rapid; @siha2006telecommuting].
Previous research has also used a variety of definitions to measure the level of
telecommuting within different workforces, distinguishing, for example, between those
directly employed, indirectly employed, self-employed, full-time or part-time, and
those who use digital technologies to work remotely full-days or part-days
[@allen2015effective; @bailey2002review; @haddad2009examination].
No matter the definition, the option and capability to telecommute or work from home
has never been equally distributed spatially or socio-economically any more than
different industries and employment opportunities have.
For example, studies from the United States, the Netherlands, and the UK indicate
that telecommuters are most likely to hold professional, managerial, and technical
occupations, where the workforce is better educated and wealthier, and that there is
suppressed demand among women and part-time workers
[@headicar2016move; @peters2004employees; @singh2013modeling].
Opportunities for working from home during the current pandemic have likewise not
been equally spread across the workforce.
@NBERw26948 indicated that in the US, managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from home,
and that occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in more spatially fixed occupations,
from farming, construction and manufacturing to hospitality and care services.
In the US, these occupations tend to be lower-income, non-white,
without a university degree, live in rental accommodation and lack health insurance
[@NBERw27085].
Similar trends can be observed for other countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, whilst only $34$% of workers with secondary education and
$14$% of those primary education did so [@eurofound2020].

## Digital divides and economic resilience
<!-- WFH and digital divides-->
Our understanding of telecommuting as a product of enabled occupations can be
described as a manifestation of the third level digital divide, as those who are able
to use digital technologies to work from home benefit from a high rate of return on
their use of the internet in terms of autonomy, flexibility, and time saved from
commuting [@peters2004employees; @siha2006telecommuting; @singh2013modeling].
These returns have been even greater during the Covid-19 crisis,
when those with the ability to telecommute also have the ability to maintain
their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the first
level digital divide, which is associated with access and quality of internet
connectivity.
For example, the systematic review from @SALEMINK2017360 highlights the
infrastructure quality differences between urban and rural areas
in various advanced economies.
Whether this variation in infrastructure quality affects the spatial footprint of
telecommuting has not previously been measured, although there are indications that
those who purchase high speed connections consume more data of all sorts and use their
connections for a variety of purposes [@hauge2011consumer].
Another study identifies a correlation between access to internet services and a
reduction in household transport spend across 33 countries [@bris2017ict].
Whether the implication is less travel because of increased telecommuting,
or whether internet access enables more efficient travel, this finding is
an indication of the potential household savings better internet services offer.
Thus, the extreme demand during the pandemic provides a new opportunity to
understand how infrastructure accessibility, quality, and reliability
affects telecommuting, particularly as working from home during the pandemic required
high volumes of bandwidth-intensive video conferencing in order to avoid the
face-to-face contact that could increase the spread of infection.
By first answering questions about internet service resilience, we can also refine our
understanding of how this has contributed to or reduced the new digital divide,
where economic resilience has been dependent upon the capability to work from home.

<!-- Covid-19 and working from home, cities, urban structure -->
The multi-layered digital divides intersect with materials divides and the economic
geography of the UK.
Following the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic],
different places have different industrial and occupational profiles,
and these affect the aggregated potential capacity of places for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south part of the UK regarding, for example, skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers some of the highest level of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
Not only all three levels of digital divides are, to a certain extend, associated
with or shaped by the geography of the UK, but the intersection of the digital and
material divides affects the capacity of places to overcome, at least partially,
the economic effects of the Covid-19 pandemic.
Importantly, this is the first time that digital technologies became an essential
tool for economic resilience for such a great part of the population.

@@

<!-- clustering in general -->
The starting point of our methodological framework is cluster analysis, which can
be defined within the modern machine learning framework as an unsupervised
learning task, which involves partitioning unlabelled observations into homogeneous
groups known as clusters [@montero2014tsclust].
The key idea is that observations within clusters tend to be more similar than
observations between clusters.
Clustering is particularly useful for exploratory studies as it identifies
structures within the data [@aghabozorgi2015time].
Therefore, cluster analysis is a widely used family of techniques in geography
[@gordon1977classification; @everitt1974cluster].
For instance, clustering methods is the basis of *geodemographics*, which aims
to create small area indicators or typologies of neighbourhoods based on various
and some times diverse variables [@SINGLETON2009289; @harris2005geodemographics].
Clustering techniques have also been employed to solve regionalisation problems
[@niesterowicz2016].

<!-- time-series clustering -->
Common characteristic of these studies is the cross-sectional nature of the data
they employ.
Indeed, most clustering problems in geography deal with observations that are
fixed in time.
However, for this paper we are interested in creating clusters of local authorities
in the UK with similar temporal signatures of experienced internet speeds over time.
Hence, we deviate from the established geographical clustering tools and
employ time-series clustering methods.

@@

Time-series clustering methods utilising the whole length of time-series can be
grouped in three categories.
The first -- model-based approaches -- is based on recovering the underlying model
for each time-series and then applying clustering algorithms on the model parameters
of each time-series [@aghabozorgi2015time].
The main criticism is the cluster accuracy for nearby clusters [@mitsa2010temporal].
The second approach is a based the formation of vectors of features based on the
original time-series.
These new data of reduced dimensionality is then clustered using conventional
clustering algorithms.

For this paper we utilise the third category of time-series clustering methods
known as shape-based approaches.
These methods match two sperate time-series objects based on the similarity of
their shapes through the calculation of distances among these two shapes and are
better equipped to capture similarities between short length
time-series [@aghabozorgi2015time].
This approach serves best this paper because (i) we aim to identify clusters of
UK local authorities with similar temporal signatures -- i.e. shapes -- of
experienced internet speeds and (ii) the length of our time-series is short
(see the data discussion in this section).

Another import element of time-series clustering is the actual clustering algorithm.
Similar to clustering of cross-sectional data, we can employ partitioning
algorithms, which lead to non-overlapping clusters, hierarchical
clustering, which result to a hierarchy of clusters and fuzzy algorithms, which
create overlapping clusters [@sardatime].
Because of the simplicity of the implementation and the interpretability of the
results, we utilise here partitioning clustering based on the widely used *k-Means*
algorithm.
This is an iterative algorithm, which begins with defining the desired number of
clusters *k*.
Then, each observation is randomly assigned to a cluster from the $[1,k]$ space.
This is the initial cluster assignment, which is followed by iterations in order
to minimise the distance between the centroids of the clusters and the observations
assigned to these clusters [@james2013introduction].

<!-- difference between clustering cross-sectional and time-series data -->
There are a number of differences between the above described application of
*k-Means* for cross-sectional data and its application for times series data.
Instead of creating clusters around centroids, a common approach is to create
clusters around *medoids*, which are representative time-series objects with a
minimal distance to all other cluster objects [@sardatime].
Also, instead of calculating the Euclidean distance between centroids and data
points, more complex distance measures need to be employed in order to capture
the similarity between a time-series object and a medoid.
A common distance measure for shape-based time-series clustering in the Dynamic
Time Warping (DTW).
Using its underpinning dynamic programming algorithm, DTW compares two time-series
objects to find the optimum warping path between them.
DTW is widely used in order to overcome limitations linked to the use of Euclidean
distance [@sardatime; @berndt1994using; @ratanamahatana2004everything].
The `R` package `dtwclust` has been used for the time-series clustering [@dtwclust].

<!-- introduction of our broadband speeds data -->
To assess the quality and reliability of internet across local authorities
in the UK during the time when the population were told to work from home if at
all possible we utilise unique data regarding individual internet speed tests from
Speedchecker Ltd^[[https://www.broadbandspeedchecker.co.uk/](https://www.broadbandspeedchecker.co.uk/)].
This a private company that allows internet users to check their own broadband
upload and download speeds, and stores every speed-check with timestamp and
geolocation information.
These data have been used before to assess digital divides [@riddlesden2014broadband]
and the impact of local loop unbundling regulatory processes [@nardotto2015unbundling].
These volunteered geographic data enable to assess the *experienced* internet
speed by users, which may differ from the *advertised* maximum speeds provided
by Internet Service Providers.
We are particularly interested in upload speeds and the frequency of speed tests.
While the former is less associated with internet-based leisure activities such
as video streaming, which involve downloading large amounts of data, the latter
can be linked more with work-related activities such as uploading documents or
two-way communication and interaction systems.
**ADD MORE. HOW ABOUT SPEED TESTS?**

<!-- data transformation -->
The first step in the workflow after dropping some outliers following
@riddlesden2014broadband is to transform and the individual,
geolocated and time-stamped tests to more meaningful aggregates both in terms
of space and time.
Regarding the temporal dimension, we aggregate all the speed-checks during the $13$
weeks of March to May inclusive for weekdays in $2020$ by each hour of the day and
day of the week.
As our research aims to identify the geography of internet service resilience for
work purposes, bank holidays and the hours between midnight and 6am were also
excluded.
The composite week time-series thus comprise 18 hours multiplied by 5 weekdays or
90 time points per series.
We aggregate these data because, we could not follow individuals or households
and connect data points.
Also, although this is a large data set ($241,088$ individual tests during the
working hours of study period), there are not enough observations for each local
and for each working hour ($631$ speed test per LAD in average).
These time-series were calculated for each of the 382 Local Authority Districts
(LADs) in the UK, standardised, and then a *k*-means partitioning around medoids
clustering algorithm was applied using DTW.
We initially run the algorithm for $k = 5, 10, 15 and 20$ and used cluster validity
indices (CVIs) to pick the optimal solution of $k = 10$.
Following @sardatime the following CVIs were employed: Silhouette (max), Score
function (max), Calinski-Harabasz (max), Davies-Bouldin (min), Modified Davies-Bouldin (DB*, min), Dunn (max), COP (min).

@@

We then match the cluster membership to the LAD-level data to identify the
characteristics of each cluster, including number of LADs, the descriptive
statistics of upload speeds in that cluster, and the temporal profile by hour of
the day and day of the composite week.
Thus we aim to answer the first part of our research question:
How resilient are broadband speeds as experienced in different parts of the UK
during a time of extreme demand?

<!-- metrics -->
Since the quality and reliability of internet services vary in time and space due
to both supply and demand-side influences, we use a number of different measures
of experienced upload speeds.
These include: a) mean, experienced connection speed,
b) standard deviation or the amount of fluctuation from the mean, and
c) the variation in speeds at particular times of day when working from home is
more likely to take place.
We take account of all three measurements in order to describe upload speeds as
fully as possible.
The following **graphs / tables** show these statistics for the sample as a whole,
as well as similar statistics related to frequency of testing.

@@

The identification strategy is as follows.
Based on the outcomes of the time-series clustering, we identify $J$ distinct
and crisp clusters.
We then regress this cluster membership against a vector $X_{i}$ of
socio-economic and geographic variables, which we are discussed in details in
**section XX**.
This analysis enables us to provide a more nuanced understanding of how
telecommuting and technology intersect at a time of extreme demand, and what
lessons this time has for a future where telecommuting is likely to remain a
common means of accessing work and broadband services, as well as infrastructure,
must be fit for purpose.

>>>>

<!-- our approach -->
Using the UK as a case study, this paper aims to understand how the quality and
reliability of internet service, as reflected in *experienced* internet speeds,
may reinforce or redress the spatial and social dimensions of the digital division
exposed by the pandemic.
To do so, we employ volunteered geographic data on individual broadband speed tests
and state-of-the-art time-series clustering methods to create clusters of UK local
authorities with similar temporal signatures of experienced internet speeds.
We then associate these clusters of local authorities with their socioeconomic
and geographic characteristics to explore how they overlap with or diverge from
the existing economic and digital geography of the UK. 
Our analysis enables us to better understand how the spatial and social distribution
of both occupations and online accessibility intersect to enable or hinder the
practice of telecommuting at a time of extreme demand.
We will also consider what lessons can be learned from this time for a future where
telecommuting is likely to remain a common means of accessing work, and broadband
services and infrastructure must be fit for purpose.
**LET'S LEAVE IT FOR NOW, BUT I THINK WE CAN CRYSTALISE MORE THE RQ**

@@

The structure of this paper is as follows.
First we review the literature on telecommuting and digital divides to better
understand the origins of the new digital divide revealed by the pandemic and its
impact on the economic resilience of different places.
We then describe our data and methodology.
Our results section first offers a classification of how internet services vary
across the UK local authorities and then assesses whether these clusters replicate
or repudiate other socio-economic and geographic patterns of economic resilience.

@@

<!-- Who works from home-->
In this analysis, the terms ‘telecommuting’ and ‘working from home’ are used
interchangeably, as most remote labour during the Covid-19 crisis was carried out
in the homes of individual employees rather than any other location [@eurofound2020].
However, it should be noted that previous research has explored how telecommuting
can occur in other places, including satellite offices or on public transport
[@felstead2012rapid; @siha2006telecommuting].
Previous research has also used a variety of definitions to measure the level of
telecommuting within different workforces, distinguishing, for example, between those
directly employed, indirectly employed, self-employed, full-time or part-time, and
those who use digital technologies to work remotely full-days or part-days
[@allen2015effective; @bailey2002review; @haddad2009examination].
No matter the definition, the option and capability to telecommute or work from home
has never been equally distributed spatially or socio-economically any more than
different industries and employment opportunities have.
For example, studies from the United States, the Netherlands, and the UK found
that telecommuters are most likely to hold professional, managerial, and technical
occupations where the workforce is better educated and wealthier, and that there is
suppressed demand among women and part-time workers
[@headicar2016move; @peters2004employees; @singh2013modeling].
Opportunities for working from home during the current pandemic have likewise not
been equally spread across the workforce.
@NBERw26948 indicated that in the US, managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from home,
and that occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in more spatially fixed occupations,
from farming, construction and manufacturing to hospitality and care services.
In the US, these occupations tend to be lower-income, non-white,
without a university degree, live in rental accommodation and lack health insurance
[@NBERw27085].
Similar trends can be observed for other countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, whilst only $34$% of workers with secondary education and
$14$% of those primary education did so [@eurofound2020].

## Digital divides and economic resilience
<!-- WFH and digital divides-->
Our understanding of telecommuting as a product of enabled occupations can be
described as a manifestation of the third level digital divide, as those who are able
to use digital technologies to work from home benefit from a high rate of return on
their use of the internet in terms of autonomy, flexibility, and time saved from
commuting [@peters2004employees; @siha2006telecommuting; @singh2013modeling].
These returns have been even greater during the Covid-19 crisis,
when those with the ability to telecommute also have the ability to maintain
their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the first
level digital divide, which is associated with access and quality of internet
connectivity at a time of extremely high demand.
@SALEMINK2017360 provides a systematic review of the pre-pandemic, 
first level digital divide in infrastructure quality between urban and rural areas 
in various advanced economies.
Yet whether this variation in infrastructure quality affects the spatial footprint of
telecommuting has not previously been measured.
There are indications that those who purchase high speed connections consume 
more data of all sorts and use their connections for a variety of purposes 
[@hauge2011consumer], 
and that there is a correlation between access to internet services and a
reduction in household transport spend [@bris2017ict].
Whether the implication is more internet use and less travel 
because of increased telecommuting, these studies suggest that better internet 
services enable households to make savings and efficiencies, 
an example of the first level digital divide reinforcing the third level.
Yet did the purchase of high speed connections and increased internet access also  
prepare households for long-term home-working, enforced by government restrictions?
The extreme demand during the pandemic provides a new opportunity to understand 
how infrastructure accessibility, quality, and reliability affects telecommuting, 
particularly in light of the high volumes of bandwidth-intensive video conferencing 
required in order to avoid the face-to-face contact that could increase 
the spread of infection. 
We seek to answer how internet service resilience contributes to or reduces 
economic resilience 
when the latter is dependent upon the capability to work from home, 
and thus improve our understanding of this new digital divide.
**RELATED TO YOUR POINT ABOVE, WE MAY CONCLUDE THIS IS NOT A NEW DIGITAL DIVIDE, BUT THAT IS STILL ANSWERING A USEFUL QUESTION - MIGHT END UP REWORDING AS INTERACTION BETWEEN DIGITAL DIVIDES?**

<!-- Covid-19 and working from home, cities, urban structure -->
The multi-layered digital divides intersect with material divides and the economic
geography of the UK.
Following the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic],
different places have different industrial and occupational profiles,
and these affect the aggregated potential capacity of places for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south part of the UK regarding, for example, skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers some of the highest level of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
Not only are all three levels of digital divides associated to a certain extent 
with or shaped by the geography of the UK, but the intersection of the digital and
material divides affects the capacity of places to overcome
some of the economic effects of the Covid-19 pandemic.
Importantly, this is the first time that digital technologies became an essential
tool for economic resilience for such a great part of the population.

@@

<!-- clustering in general -->
The starting point of our methodological framework is cluster analysis, which can
be defined within the modern machine learning framework as an unsupervised
learning task, partitioning unlabelled observations into homogeneous groups 
known as clusters [@montero2014tsclust].
The key idea is that observations within clusters tend to be more similar than
observations between clusters.
Clustering is particularly useful for exploratory studies as it identifies
structures within the data [@aghabozorgi2015time].
Therefore, cluster analysis is a widely used family of techniques in geography
[@gordon1977classification; @everitt1974cluster].
For instance, clustering methods are the basis of *geodemographics*, a discipline 
which aims to create small area indicators or typologies of neighbourhoods 
based on various and sometimes diverse variables 
[@SINGLETON2009289; @harris2005geodemographics].
Clustering techniques have also been employed to solve regionalisation problems
[@niesterowicz2016].

<!-- time-series clustering -->
Common characteristics of these studies are the cross-sectional nature of the data
they employ.
Indeed, most clustering problems in geography deal with observations that are
fixed in time.
However, for this paper we are interested in internet speeds, which vary over time. 
We aim to create clusters of local authorities in the UK with similar temporal 
signatures of experienced internet speeds.
Hence, we deviate from the established geographical clustering tools and
employ time-series clustering methods.

@@

Time-series clustering methods utilising the whole length of time-series can be
grouped in three categories.
The first -- model-based approaches -- is based on recovering the underlying model
for each time-series and then applying clustering algorithms on the model parameters
of each time-series [@aghabozorgi2015time].
The main criticism is the cluster accuracy for nearby clusters [@mitsa2010temporal].
The second approach is based the formation of vectors of features derived from the
original time-series.
These new data of reduced dimensionality are then clustered using conventional
clustering algorithms.

For this paper we utilise the third category of time-series clustering methods
known as shape-based approaches.
These methods match two sperate time-series objects based on the similarity of
their shapes through the calculation of distances between the shapes, 
and are thus better equipped to capture similarities between short length
time-series [@aghabozorgi2015time].
This approach serves best this paper because (i) we aim to identify clusters of
UK local authorities with similar temporal signatures -- i.e. shapes -- of
experienced internet speeds and (ii) the length of our time-series is short
(see the data discussion in this section).

Another import element of time-series clustering is the actual clustering algorithm.
Similar to the clustering of cross-sectional data, we can employ partitioning
algorithms, which lead to non-overlapping clusters, hierarchical clustering, 
which classifies clusters at different levels, and fuzzy algorithms, which
create overlapping clusters [@sardatime].
Because of the simplicity of the implementation and the interpretability of the
results, we utilise here partitioning clustering based on the widely used *k-Means*
algorithm.
This is an iterative algorithm, which begins with defining the desired number of
clusters *k*.
Then each observation is randomly assigned to a cluster from the $[1,k]$ space.
This initial cluster assignment is followed by iterations in order to minimise 
the distance between the centroids of the clusters and the observations
assigned to these clusters [@james2013introduction].

<!-- difference between clustering cross-sectional and time-series data -->
There are a number of differences between the above described application of
*k-Means* for cross-sectional data and its application for times series data.
Instead of creating clusters around centroids, a common approach is to create
clusters around *medoids*, which are representative time-series objects with a
minimal distance to all other cluster objects [@sardatime].
Also, instead of calculating the Euclidean distance between centroids and data
points, more complex distance measures need to be employed in order to capture
the similarity between a time-series object and a medoid.
A common distance measure for shape-based time-series clustering is 
Dynamic Time Warping (DTW).
Using its underpinning dynamic programming algorithm, DTW compares two time-series
objects to find the optimum warping path between them.
DTW is widely used in order to overcome limitations linked to the use of Euclidean
distance [@sardatime; @berndt1994using; @ratanamahatana2004everything].
The `R` package `dtwclust` has been used for the time-series clustering [@dtwclust].

<!-- introduction of our broadband speeds data -->
To assess the quality and reliability of internet across local authorities
in the UK during the time when the population were told to work from home if at
all possible we utilise unique data comprising individual internet speed tests from
Speedchecker Ltd^[[https://www.broadbandspeedchecker.co.uk/](https://www.broadbandspeedchecker.co.uk/)].
This a private company that allows internet users to check their own broadband
upload and download speeds, and stores every speed-check with timestamp and
geolocation information.
These data have been used before to assess digital divides [@riddlesden2014broadband]
and the impact of local loop unbundling regulatory processes [@nardotto2015unbundling].
These volunteered geographic data enable to assess the *experienced* internet
speed by users, which may differ from the *advertised* maximum speeds provided
by Internet Service Providers.
We are particularly interested in upload speeds and the frequency of speed tests.
Upload speeds are less associated with internet-based leisure activities such
as video streaming, which involve downloading large amounts of data, but they
can be linked more with work-related activities such as uploading documents or
two-way communication and interaction systems.
Meanwhile, the frequency of speed tests was important in identifying 
the temporal profile which woud give us most insight into 
experienced internet service and resilience.
Although we know from government statements the start and end of the period of
interest, namely, March to May 2020, the frequency of speed tests within that period 
provides an indication of the volume of experience over any particular units of time.

<!-- data transformation -->
Thus, the first step in the workflow after dropping some outliers following
@riddlesden2014broadband was to transform the individual, geolocated and time-stamped 
tests to more meaningful aggregates both in terms of space and time.
The frequency of testing indicates that whilst there is an overall trend of increased
testing or increased experience of internet service from March to April and then a 
slight reduction from April to May, this trend masks substantial variation 
by not only the day of the week, but also time of day, as can be seen in **[graph]**.
Thus, a daily aggregation of upload speeds would also mask this pattern.
Furthermore, the importance of this pattern is highlighted by a comparison 
with the same period in 2019 **see graph**, 
when the volume of testing and thus of experience of internet service quality
was much more tied to the evening leisure peak and presumably to download speeds.
Since the increase in testing during the working day in 2020 is an indication that 
users have a greater perception of the variable quality of internet service,
particularly during a new morning peak of testing for service reliability,
we decided to include a measure of hourly variation in our temporal profiles.
However, although this is a large data set ($241,088$ individual tests during the
working hours of study period), there are not enough observations for each
Local Authority District (LAD) and for each working hour of each working day
($631$ speed tests per LAD on average) to profile speeds at that level of detail.
Therefore, we aggregate all the speed-checks during the $13$ weeks of March to May 
inclusive for weekdays in $2020$ by each hour of the day and day of the week.
As our research aims to identify the geography of internet service resilience for
work purposes, bank holidays and the hours between midnight and 6am were excluded,
as well as weekend days.
The composite week time-series thus comprise 18 hours multiplied by 5 weekdays or
90 time points per series.
We also aggregate these data spatially because we could not follow individuals or 
households and connect data points.
The time-series were calculated for each of the 382 LADs in the UK, standardised, 
and then a *k*-means partitioning around medoids clustering algorithm 
was applied using DTW.
We initially run the algorithm for $k = 5, 10, 15 and 20$ and used cluster validity
indices (CVIs) to pick the optimal solution of $k = 10$.
Following @sardatime the following CVIs were employed: Silhouette (max), Score
function (max), Calinski-Harabasz (max), Davies-Bouldin (min), Modified Davies-Bouldin (DB*, min), Dunn (max), COP (min).

@@

<!-- metrics -->
We then review the characteristics of each cluster, including number of LADs, 
and the temporal profile of upload speed by hour of the day 
and day of the composite week.
Since the quality and reliability of internet services vary in time and space due
to both supply and demand-side influences, we use a number of different measures
to then describe experienced upload speeds per cluster.
These include: a) mean, experienced connection speed,
b) standard deviation or the amount of fluctuation from the mean, and
c) the variation in speeds at particular times of day when working from home is
more likely to take place.
We take account of all three measurements in our descriptives statistics of
upload speeds in order to determine how resilient broadband speeds are 
as experienced in different parts of the UK during a time of extreme demand.

@@

The identification strategy is as follows.
Based on the outcomes of the time-series clustering, we identify $J$ distinct
and crisp clusters.
We then regress this cluster membership against a vector $X_{i}$ of
socio-economic and geographic variables, which are discussed in detail in
**section XX**.
This analysis enables us to provide a more nuanced understanding of how
telecommuting and technology intersect at a time of extreme demand, and what
lessons this time has for a future where telecommuting is likely to remain a
common means of accessing work and broadband services, as well as infrastructure,
must be fit for purpose.
